Bayesian Rate Estimation

### 2012-08-16 many.tex
Finished section on loudest event and foreground dominated statistics. Added figure for GW example.

\newcommand{\ilya}[1]{{\color{red} \bf #1}}
\newcommand{\will}[1]{{\color{blue} \bf #1}}
\newcommand{\jon}[1]{{\color{orange} \bf  #1}}

@@

We assume that we are presented with a data set of $N$ events that exceed a pre-specified threshold in ranking statistic, $x_{min}$.  Each
event may be due to either a signal of interest or an uninteresting
background.  Each event is associated with a ranking statistic, $x$.
Our data set therefore consists of the ranking statistics for the set
of events:
\begin{equation}
  d = \mathset{ x_i | i = 1, \ldots, N } .
\end{equation}
The number of events $N$ is also part of the observed data, but we separate out $N$ and the observed ranking statistics, $d$, for convenience. We can choose how to label our events. Ultimately we will label the events in order of ranking statistic, i.e., $x_1 < x_2 < \cdots < x_N$, but some of the derivations that follow are simpler if the events are ordered by time of arrival. We will use $d$ to denote ranking statistic-ordered events, and $d_{\rm to}$ to denote time-ordered events.

@@

For the usual $k=1$ case, the marginalised posterior (Eq.~\eqref{eq:rate-shape-posterior}) becomes
\begin{equation}
p_{\rm LE}\left( R_f, R_b, \theta | d \right) \propto \left(R_f \hat{f}(x_N,\theta) + R_b \hat{b}(x_N,\theta)\right)\exp\left[-\left( R_f (1-\hat{F}(x_N,\theta)) +
      R_b (1-\hat{B}(x_N,\theta)) \right) \right] \frac{p(\theta)}{\sqrt{R_f R_b}}.
\end{equation}
where $x_N$ denotes the loudness of the loudest event. Marginalising over $R_b$ we obtain
\begin{equation}
p_{\rm LE}\left( R_f, \theta | d \right) \propto \left( \frac{\hat{b}(x_N,\theta)}{2(1-\hat{B}(x_N,\theta))} + R_f \hat{f}(x_N,\theta)\right)\frac{\sqrt{\pi}}{\sqrt{1-\hat{B}(x_N,\theta)}\,\sqrt{R_f}} \exp(-R_f(1-\hat{F}(x_N,\theta))) .
\end{equation}
This posterior has a maximum in $R_f$ at
\begin{equation}
R_f = \frac{\hat{f}(x_N,\theta)-(1-\hat{F}(x_N,\theta)\tilde{b}(x_N,\theta) \pm \sqrt{(\hat{f}(x_N,\theta)-(1-\hat{F}(x_N,\theta)\tilde{b}(x_N,\theta))^2 - 4\tilde{b}(x_N,\theta)(1-\hat{F}(x_N,\theta))\hat{f}(x_N,\theta)}}{4 \hat{f}(x_N,\theta)(1-\hat{F}(x_N,\theta)}.
\end{equation}
where $\tilde{b}(x_N,\theta) =\hat{b}(x_N,\theta)/(1-\hat{B}(x_N,\theta))$. If $\tilde{b}(x_N,\theta) \ll 1$, i.e., the loudest event is very likely to be foreground, we obtain the result $(1- \hat{F}(x_N,\theta))\,R_f \approx 1/2$. This is can be understood as the statement that the rate of foreground events with ranking statistic greater than $x_N$,  $(1- \hat{F}(x_N,\theta))\,R_f$, is of order $1$, as expected. If $\tilde{b}(x_N,\theta) >> 1$, however, the posterior on $R_f$ is peaked at $0$.

\subsubsection{Foreground dominated statistic}
If we set the threshold for including an event, $x_{\rm min}$, sufficiently high, we can ensure that $\hat{f}(x_i,\theta) \gg \hat{b}(x_i,\theta)$ for all ranking statistics $x_i$ in the data set. The posterior can then be approximated by
\begin{equation}
p_{\rm FD}\left( R_f, R_b, \theta | d \right) \propto  \prod_{i} \left[ \hat{f}\left(x_i, \theta\right)\right] R_f^N \exp\left[-\left( R_f + R_b \right) \right] \frac{p(\theta)}{\sqrt{R_f R_b}}.
\end{equation}
Normalisation over $R_b$ gives a constant factor and the posterior on the foreground rate becomes
\begin{equation}
p_{\rm FD}\left( R_f, \theta | d \right) \propto  \prod_{i} \left[ \hat{f}\left(x_i, \theta\right)\right] R_f^{N-\frac{1}{2}} \exp\left[-R_f  \right] p(\theta).
\end{equation}
Ignoring the dependence on $\theta$, this is peaked at a rate $R_f = N-1/2$, so we have the expected result that, in the foreground dominated regime, the rate is approximately equal to the number of events observed. We note that this is the rate of events occurring above threshold and so if we were to use this statistic with different choices of threshold we would have to renormalise the rates to a fixed threshold in order to compare the results.

@@

Suppose we attempt to detect gravitational wave signals in a data
stream by matched filtering in the frequency domain against a set of
$N$ template waveforms \citep[e.g.,][]{findchirppaper,LVC2011}.  In our simplistic
model, we suppose the data stream consists of stationary Gaussian
noise with a power spectral density $S(f)$ combined additively with
some number of gravitational wave signals.  We assume that the signals
are sufficiently rare that they do not overlap in the data stream.
The signal to noise ratio (SNR) of a template, $h(f)$, given data,
$d(f)$, is
\begin{equation}
  \rho_h \equiv \frac{\left\langle h, d \right\rangle}{\sqrt{\left
        \langle h, h \right\rangle}},
\end{equation}
where $\left \langle \cdot \right\rangle$ denotes the noise-weighted
inner product:
\begin{equation}
  \left\langle a, b \right\rangle \equiv \ilya{4 \Re} \int_0^\infty df\,
  \frac{a^*(f) b(f)}{S(f)}.
\end{equation}
We suppose for simplicity that the templates are sufficiently distinct
that 
\begin{equation}
  \left\langle h_i, h_j \right\rangle \simeq \delta_{ij}.
\end{equation}
In the following subsection, we will generalize the model to
overlapping templates.  We rank candidate events by their maxmim SNR
over the entire template bank,
\begin{equation}
  x \equiv \max_{h} \rho_h,
\end{equation}
and consider only events that have a maximum SNR above some threshold,
$x > \xmin$.

For a data stream of pure noise, $d(f) = n(f)$, the SNR of a each
template follows a $N(0,1)$ distribution.  The background ranking
statistic (i.e.\ the maximum SNR over the template bank) then has a
cumulative distribution without thresholding of
\begin{equation}
  \hat{B}(x) = \left( \frac{1 + \erf\left( \frac{x}{\ilya{\sqrt{2}}}
      \right)}{2} \right)^N
\end{equation}
Imposing the threshold, $x > \xmin$, the
cumulative distribution of the background becomes
\begin{equation}
  \label{eq:analytic-background-rate}
  \hat{B}(x) = \frac{\left( 1 + \erf\left( \frac{x}{\ilya{\sqrt{2}}} \right)
    \right)^N - \left( 1 + \erf\left( \frac{\xmin}{\ilya{\sqrt{2}}} \right)
    \right)^N}{2^N - \left( 1 + \erf\left( \frac{\xmin}{\ilya{\sqrt{2}}} \right)
    \right)^N }
\end{equation}
\ilya{for $x>\xmin$, $0$ otherwise.}

@@

\ilya{[General comment on the text: I think we need to say a bit more
    about the figures and the interpretation of the results, instead
    of just relying on the reader to make the best of it.  For
    example, even though it may seem obvious to us, we should
    explicitly say that Fig.~2 shows that we succeed in recovering
    rates within the expected uncertainty envelope.  On Fig.~3, could
    say something obvious like point out that we cannot separate
    background from foreground for individual events, but that doesn't
    necessarily matter for rate measurements.  Now, one interesting
    question that Fig.~3 brings to mind is the following: if we only
    used "gold-plated" events (say, those with
    $p(x|fore)/p(x|back)>0.99$, how much worse / more uncertain would
    the rate estimation be?]}

@@

@@

In \S \ref{sec:analytic-GW-example} we assumed that the overlap
between \ilya{different} templates in the template bank \ilya{was
  negligible}, so the SNRs \ilya{recovered by} different templates are
independent random variables.  In fact, template banks are not
constructed in this way \citep[e.g.,][]{Owen:1998dk,Ajith:2008},
%\citep{Caudill2011,Cannon2011}, 
because signals could fall in the gaps between the non-overlapping
templates.  We can model this effect by assuming that a template bank
of $N$ actual templates will behave as if it had $N_\mathrm{eff}$
\emph{independent} templates.  \ilya{Rather than pre-computing
  $N_\mathrm{eff}$,} we can fit for \ilya{it as a} shape parameter.
That is, we assume that \ilya{$\theta = \{N_\mathrm{eff}\}$} is a
shape parameter for the background cumulative distribution:
\begin{equation}
  \hat{B}\left(x, N_\mathrm{eff}\right) = \frac{\left( 1 + \erf\left( \frac{x}{2} \right)
    \right)^{N_\mathrm{eff}} - \left( 1 + \erf\left( \frac{\xmin}{2} \right)
    \right)^{N_\mathrm{eff}}}{2^{N_\mathrm{eff}} - \left( 1 + \erf\left( \frac{\xmin}{2} \right)
    \right)^{N_\mathrm{eff}} }.
\end{equation}

@@

\ilya{[Again, I think we need to explicitly compare this method with
    two alternatives.  Alternative one is the loudest event statistic.
    Alternative two is just counting gold-plated events, and dividing
    the volume in which we are sensitive to them by their number.  In
    principle, alternative one should be suboptimal when the number of
    events is more than 0 or 1, while alternative two should be OK
    when the number of events is large, but error-prone otherwise.
    But if we want people to use this technique, we have to defend the
    point that there is a regime where it is clearly superior (and say
    what that regime is).]}

>>>>

\newcommand{\ilya}[1]{{\color{red} \bf #1}}
\newcommand{\will}[1]{{\color{blue} \bf #1}}
\newcommand{\jon}[1]{{\color{green} \bf  #1}}

@@

We assume that we are presented with a data set of $N$ events that exceed a pre-specified threshold in ranking statistic, $x_{\rm min}$.  Each
event may be due to either a signal of interest or an uninteresting
background.  Each event is associated with a ranking statistic, $x$.
Our data set therefore consists of the ranking statistics for the set
of events:
\begin{equation}
  d = \mathset{ x_i | i = 1, \ldots, N } .
\end{equation}
The number of events $N$ is also part of the observed data, but we separate out $N$ and the observed ranking statistics, $d$, for convenience. We can choose how to label our events. Ultimately we will label the events in order of ranking statistic, i.e., $x_1 < x_2 < \cdots < x_N$, but some of the derivations that follow are simpler if the events are ordered by time of arrival. We will use $d$ to denote ranking statistic-ordered events, and $d_{\rm to}$ to denote time-ordered events.

@@

For the usual $k=1$ case, the marginalised posterior (Eq.~\eqref{eq:rate-shape-posterior}) becomes
\begin{equation}
p_{\rm LE}\left( R_f, R_b, \theta | d \right) \propto \left(R_f \hat{f}(x_N,\theta) + R_b \hat{b}(x_N,\theta)\right)\exp\left[-\left( R_f (1-\hat{F}(x_N,\theta)) +
      R_b (1-\hat{B}(x_N,\theta)) \right) \right] \frac{p(\theta)}{\sqrt{R_f R_b}}.
\end{equation}
where $x_N$ denotes the loudness of the loudest event. Marginalising over $R_b$ we obtain
\begin{equation}
p_{\rm LE}\left( R_f, \theta | d \right) \propto \left( \frac{\hat{b}(x_N,\theta)}{2(1-\hat{B}(x_N,\theta))} + R_f \hat{f}(x_N,\theta)\right)\frac{\sqrt{\pi}\, p(\theta)}{\sqrt{1-\hat{B}(x_N,\theta)}\,\sqrt{R_f}} \exp\left(-R_f(1-\hat{F}(x_N,\theta))\right) .
\end{equation}
This posterior has a maximum in $R_f$ at
\begin{equation}
R_f = \frac{\hat{f}(x_N,\theta)-(1-\hat{F}(x_N,\theta)\tilde{b}(x_N,\theta) \pm \sqrt{(\hat{f}(x_N,\theta)-(1-\hat{F}(x_N,\theta)\tilde{b}(x_N,\theta))^2 - 4\tilde{b}(x_N,\theta)(1-\hat{F}(x_N,\theta))\hat{f}(x_N,\theta)}}{4 \hat{f}(x_N,\theta)(1-\hat{F}(x_N,\theta)}.
\end{equation}
where $\tilde{b}(x_N,\theta) =\hat{b}(x_N,\theta)/(1-\hat{B}(x_N,\theta))$. If $\tilde{b}(x_N,\theta) \ll 1$, we obtain the result $(1- \hat{F}(x_N,\theta))\,R_f \approx 1/2$. This is can be understood as the statement that the rate of foreground events with ranking statistic greater than $x_N$,  $(1- \hat{F}(x_N,\theta))\,R_f$, is of order $1$, as expected. However, $\tilde{b}(x_N,\theta) = -{\rm d} [\ln(1-\hat{B}(x,\theta))]/{\rm d}x$ and $(1-\hat{B}(x,\theta)) \rightarrow 0$ as $x \rightarrow \infty$, so this term will be divergent and for many reasonable examples, we will find $\tilde{b}(x_N,\theta) >> \hat{f}(x_N,\theta)$, in which case the posterior on $R_f$ is peaked at $0$. This issue highlights the problem with using a loudest-event statistic with an improper prior on the background rate $R_b$. No matter how improbable an event with $x=x_N$ is under the background distribution, it can become likely that the event at $x_N$ is from the background distribution by taking the background rate to be sufficiently large. Although this predicts many more events with $x < x_N$, by using only the loudest event we do not incorporate the information that no such events are seen. This problem is avoided in the new framework described here, since we use all events detected above threshold and combined rates, $R_f+R_b$, significantly greater than the total number of observed events are strongly disfavoured.

This problem can be avoided in the loudest event framework, by including an upper limit on the rate, $R_{\rm max}$, in the prior for $R_b$. The marginalised distribution for the foreground rate then becomes
\begin{eqnarray}
p_{\rm LE}\left( R_f, \theta | d \right) &\propto& \left( \frac{\hat{b}(x_N,\theta)}{(1-\hat{B}(x_N,\theta))^{\frac{3}{2}}} \left[\frac{\sqrt{\pi}}{2}{\rm erf}\left(\sqrt{(1-\hat{B}(x_N,\theta))\,R_{\rm max}}\right) - \sqrt{(1-\hat{B}(x_N,\theta))\,R_{\rm max}} \, {\rm e}^{-(1-\hat{B}(x_N,\theta)) \, R_{\rm max}}\right] \right. \nonumber \\
&& \left.+ R_f \hat{f}(x_N,\theta)\frac{\sqrt{\pi}\,{\rm erf}\left(\sqrt{(1-\hat{B}(x_N,\theta))\,R_{\rm max}}\right)}{\sqrt{(1-\hat{B}(x_N,\theta))}}\right)\frac{p(\theta)}{\sqrt{R_f}} \exp\left(-R_f(1-\hat{F}(x_N,\theta))\right) \label{LEwithRmax},
\end{eqnarray}
where ${\rm erf}(x)$ is the error function, defined in the usual way ${\rm erf}(x) = (2/\sqrt{\pi}) \int_0^x \exp(-u^2){\rm d}u$. If $(1-\hat{B}(x_N,\theta)) \, R_{\rm max} \ll 1$, Eq.~(\ref{LEwithRmax}) can be approximated by
\begin{equation}
p_{\rm LE}\left( R_f, \theta | d \right) \propto \left( \frac{R_{\rm max}}{3} \hat{b}(x_N,\theta) + R_f \hat{f}(x_N,\theta)\right)\frac{p(\theta)}{\sqrt{R_f}} \exp\left(-R_f(1-\hat{F}(x_N,\theta))\right)
\end{equation}
and if $\hat{f}(x_N,\theta) \gg R_{\rm max} \hat{b}(x_N.\theta)$ we find the same result as before $(1- \hat{F}(x_N,\theta))\,R_f \approx 1/2$.

\subsubsection{Foreground dominated statistic}
If we set the threshold for including an event, $x_{\rm min}$, sufficiently high, we can ensure that $\hat{f}(x_i,\theta) \gg \hat{b}(x_i,\theta)$ for all ranking statistics $x_i$ in the data set. The posterior can then be approximated by
\begin{equation}
p_{\rm FD}\left( R_f, R_b, \theta | d \right) \propto  \prod_{i} \left[ \hat{f}\left(x_i, \theta\right)\right] R_f^N \exp\left[-\left( R_f + R_b \right) \right] \frac{p(\theta)}{\sqrt{R_f R_b}}.
\end{equation}
Normalisation over $R_b$ gives a constant factor and the posterior on the foreground rate becomes
\begin{equation}
p_{\rm FD}\left( R_f, \theta | d \right) \propto  \prod_{i} \left[ \hat{f}\left(x_i, \theta\right)\right] R_f^{N-\frac{1}{2}} \exp\left[-R_f  \right] p(\theta).
\end{equation}
Ignoring the dependence on $\theta$, this is peaked at a rate $R_f = N-1/2$, so we have the expected result that, in the foreground dominated regime, the rate is approximately equal to the number of events observed. We note that the rate $R_f$ is defined as the rate of events occurring above the specified threshold and therefore if the foreground dominated statistic were used to estimated rates $R_f^1$ and $R_f^2$ for thresholds $x_{\rm min}=x_1$, $x_{\rm min}=x_2$ respectively, these rates should be compared by equating $R_f^1 (1-\hat{F}(x_2,\theta))$ and $R_f^2 (1-\hat{F}(x_1,\theta))$.

@@

Suppose we attempt to detect gravitational wave signals in a data
stream by matched filtering in the frequency domain against a set of
$N$ template waveforms \citep[e.g.,][]{findchirppaper,LVC2011}.  In our simplistic
model, we suppose the data stream consists of stationary Gaussian
noise with a power spectral density $S(f)$ combined additively with
some number of gravitational wave signals.  We assume that the signals
are sufficiently rare that they do not overlap in the data stream.
The signal to noise ratio (SNR) of a template, $h(f)$, given data,
$d(f)$, is
\begin{equation}
  \rho_h \equiv \frac{\left\langle h, d \right\rangle}{\sqrt{\left
        \langle h, h \right\rangle}},
\end{equation}
where $\left \langle \cdot \right\rangle$ denotes the noise-weighted
inner product:
\begin{equation}
  \left\langle a, b \right\rangle \equiv 4 \Re \int_0^\infty df\,
  \frac{a^*(f) b(f)}{S(f)}.
\end{equation}
We suppose for simplicity that the templates are sufficiently distinct
that 
\begin{equation}
  \left\langle h_i, h_j \right\rangle \simeq \delta_{ij}.
\end{equation}
In the following subsection, we will generalize the model to
overlapping templates.  We rank candidate events by their maxmim SNR
over the entire template bank,
\begin{equation}
  x \equiv \max_{h} \rho_h,
\end{equation}
and consider only events that have a maximum SNR above some threshold,
$x > \xmin$.

For a data stream of pure noise, $d(f) = n(f)$, the SNR of a each
template follows a $N(0,1)$ distribution.  The background ranking
statistic (i.e.\ the maximum SNR over the template bank) then has a
cumulative distribution without thresholding of
\begin{equation}
  \hat{B}(x) = \left( \frac{1 + \erf\left( \frac{x}{\sqrt{2}}
      \right)}{2} \right)^N
\end{equation}
where ${\rm erf}(x)$ is the error function as before. Imposing the threshold, $x > \xmin$, the
cumulative distribution of the background becomes
\begin{equation}
  \label{eq:analytic-background-rate}
  \hat{B}(x) = \frac{\left( 1 + \erf\left( \frac{x}{{\sqrt{2}}} \right)
    \right)^N - \left( 1 + \erf\left( \frac{\xmin}{{\sqrt{2}}} \right)
    \right)^N}{2^N - \left( 1 + \erf\left( \frac{\xmin}{{\sqrt{2}}} \right)
    \right)^N }
\end{equation}
for $x>\xmin$, $0$ otherwise.

@@

We can compare these results to results obtained using the two approximations described earlier, the loudest-event statistic and the foreground-dominated statistic. The marginalised distribution for the foreground rate using these alternatives are shown in Figure~\ref{fig:altmeth}. In this case, the loudest event had $x_N=9.5$\jon{(correct this)}. The loudest-event statistic depends on a specification of the maximum, $R_{\rm max}$, for the background rate. We show results for $R_{\rm max}=\infty$, i.e.m, the improper prior, and $R_{\rm max}=10000$. The results for other reasonable  choices of $R_{\rm max} = 100, 1000, 100000$ etc. gave exactly the same posterior, since $\hat{b}(x_N,) R_{\rm max} \ll \hat{f}(x_N)$ for all these choices and we are therefore in the regime where the posterior is insensitive to $R_{\rm max}$. To apply the foreground-dominated statistic we must specify a threshold above which we assume all events are foreground. It is reasonable to do this based on a specification for the relative probability of an event being fore/background, $\hat{f}(x)/\hat{b}(x)=p_{\rm thresh}$. Setting $p_{\rm thresh}=0.99$ gives $x_{\rm min}=4.07$ and there are $N=13$ events exceeding that threshold. Setting $p_{\rm thresh}=0.5$ gives $x_{\rm min}=3.82$ and there are $N=15$ events exceeding that threshold \jon{(correct these numbers using Will's data)}. We show results for both of these choices of $x_{\rm min}$ in Figure~\ref{fig:altmeth}. We also show similar results for two extreme cases --- setting $x_{\rm min}=3.5$ so that we assume all $98$ observed events are foreground; and setting $x_{\rm min}=9$ so that only the loudest event is classified as foreground.

The loudest event statistic with the improper-prior gives, as expected, a poor approximation to the foreground rate. The peak is more accurately located when a prior maximum rate is defined, but the distribution is much wider than using the full analysis described here. This is to be expected as much of the information is being thrown away. The foreground-dominated statistic gives a reasonable approximation to the true foreground rate, and a distribution that is not much wider than the full analysis, for the non-extreme choices of $x_{\rm min}$. For $x_{\rm min}=9$, the foreground dominated statistic has equivalent performance to the loudest event statistic, as we might expect, while when $x_{\rm min}=3.5$, it performs poorly since we are approximating the foreground rate by the total foreground plus background rate. This indicates that, provided the threshold is chosen appropriately, the foreground dominated statistic can perform quite well at estimating the rate. The fact that it reproduces the posterior from the full analysis so well is indicative of the fact that most of the information about the foreground comes from the loudest events. Indeed, for $x_{\rm min} \approx 4$, the set of events consists mostly of foreground sources and that set contains a decent number of events ($13$), so it is not surprising that this statistic can estimate the foreground rate well. The choice of $x_{\rm min}$ was motivated by knowledge of $\hat{b}(x,\theta)$ and $\hat{f}(x,\theta)$, which we have in this case, but this might not be possible if the parameters $\theta$ were simultaneously being estimated. The full analysis naturally incorporates inference about the rate parameters $\theta$ and the background rate $R_b$ along with the foreground rate and incorporates maximum information from the data set and should therefore lead to narrower posteriors in general.

\ilya{[General comment on the text: I think we need to say a bit more
    about the figures and the interpretation of the results, instead
    of just relying on the reader to make the best of it.  For
    example, even though it may seem obvious to us, we should
    explicitly say that Fig.~2 shows that we succeed in recovering
    rates within the expected uncertainty envelope.  On Fig.~3, could
    say something obvious like point out that we cannot separate
    background from foreground for individual events, but that doesn't
    necessarily matter for rate measurements.}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{AltMethodComp}
  \caption{\label{fig:altmeth}Posteriors on foreground rate obtained using alternative methods --- the loudest event statistic and the foreground dominated statistic. Results are shown for two different choices of the maximum a priori background rate, $R_{\rm max}$, in the case of the loudest event (``LE'') statistic, and for several different choices of the threshold, $x_{\rm min}$, in the case of the foreground dominated (``FD'') statistic.} 
\end{figure}

@@

In \S \ref{sec:analytic-GW-example} we assumed that the overlap
between different templates in the template bank was
  negligible, so the SNRs recovered by different templates are
independent random variables.  In fact, template banks are not
constructed in this way \citep[e.g.,][]{Owen:1998dk,Ajith:2008},
%\citep{Caudill2011,Cannon2011}, 
because signals could fall in the gaps between the non-overlapping
templates.  We can model this effect by assuming that a template bank
of $N$ actual templates will behave as if it had $N_\mathrm{eff}$
\emph{independent} templates.   Rather than pre-computing
  $N_\mathrm{eff}$, we can fit for it as a shape parameter.
That is, we assume that $\theta = \{N_\mathrm{eff}\}$ is a
shape parameter for the background cumulative distribution:
\begin{equation}
  \hat{B}\left(x, N_\mathrm{eff}\right) = \frac{\left( 1 + \erf\left( \frac{x}{\sqrt{2}} \right)
    \right)^{N_\mathrm{eff}} - \left( 1 + \erf\left( \frac{\xmin}{\sqrt{2}} \right)
    \right)^{N_\mathrm{eff}}}{2^{N_\mathrm{eff}} - \left( 1 + \erf\left( \frac{\xmin}{\sqrt{2}} \right)
    \right)^{N_\mathrm{eff}} }.
\end{equation}

@@

If we consider the two alternative methods, the loudest event and foreground dominated statistics, we will recover the same foreground distributions as in the case with $N_{\rm eff}$ fixed, which are shown in Figure~\ref{fig:altmeth}. This is because the parameter $N_{\rm eff}$ affects only the background distribution, to which the foreground-dominated statistic is insensitive, and in the loudest event case, after marginalisation over $N_{\rm eff}$ we find $\int_0^{N_{\rm max}} \hat{b}(x_N,N_{\rm eff}) {\rm d}N_{\rm eff} \ll 3 N_{\rm max}/R_{\rm max}  \hat{f}(x_N, N_{\rm eff})$ and so we are still in the foreground-dominated regime in which the loudest event tells us nothing about the background. Similarly, neither of these alternative methods can inform us about the value of $N_{\rm eff}$, a property of the background. \jon{Technically, if we allowed really large values of $N_{\rm eff}$ we would get into a regime where the loudest event could be background and we'd get something here. But, we would need $N_{\rm max} \sim 10^{20}$ and so I don't think we should consider these alternatives further in this case.}

%\ilya{[Again, I think we need to explicitly compare this method with
    %two alternatives.  Alternative one is the loudest event statistic.
    %Alternative two is just counting gold-plated events, and dividing
    %the volume in which we are sensitive to them by their number.  In
    %principle, alternative one should be suboptimal when the number of
    %events is more than 0 or 1, while alternative two should be OK
    %when the number of events is large, but error-prone otherwise.
    %But if we want people to use this technique, we have to defend the
    %point that there is a regime where it is clearly superior (and say
    %what that regime is).]}
